Better Corpus Visualization through Topic Ranking

### 2010-10-21 masters_thesis/topic_ranking_nips/paper.tex
Renamed the directories to make a bit more sense

\begin{abstract}
\end{abstract}

@@

The rest of this paper is outlined as follows.  We first give further
motivation and a more detailed description of the task of ranking topics by
relevance in \secref{ranking}.  \secref{docdist} describes a method for
determining the relevance of a topic to various possible questions a person
could have when looking at a corpus, and \secref{examples} shows examples of
how this technique provides a proper ranking.  In \secref{challenges} we
discuss the challenges of evaluating this visualization method.  Finally, in
\secref{conclusion} we conclude.

@@

Our use of the document distribution for a topic focuses primarily on
calculating the entropy of the distribution and of related distributions where
documents are grouped by category.  For instance, one can group the documents
by author and compute the distribution of the topic over authors as
\begin{align}
  p(a|z) = \frac{\sum_{d \in \mathcal{A}} C(d,z)}{\sum_{d \in \mathcal{D}}
  C(d,z)}
\end{align}
where $a$ is an author and $\mathcal{A}$ is the set of documents written by
that author.

@@

\section{Using $p(d|z)$ to rank topics}
\label{sec:examples}

We present here a few examples of how the document distribution for a topic can
be used to rank topics according to their relevance for certain user questions.
In our examples we use a corpus of 480 campaign speeches from the 2008
presidential primary and general election.  We inferred 150 topics in the data
from a Gibbs sampling of the LDA topic model.

Questions here:

What topics are most prevelant in the corpus?

What topics did John McCain talk about the most?

What topics did Barack Obama most consistently throughout his campaign?

What topics are most consistent across candidates?

What topics were constant throughout the campaign?

What topics were transitory, only spoken of rarely?

\section{Challenges of evaluation}
\label{sec:challenges}

The obvious way to evaluate a topic ranking is to use the evaluations methods
commonly found in information retrieval tasks.  There are well established
techniques to evaluate a ranking given a true ranking.  However, there are
several obstacles to using such a scheme to evaluate performance on this task.

The first and glaring obstacle is that there is no set of data that has been
labeled with true rankings for these or any questions.  One cannot compare the
output of an algorithm to ground truth when there is no ground truth.  It is
not even clear how such data would be obtained; perhaps one could take a topic
model and a set of questions and rank each topic for relevance to each
question, but such a task would be incredibly laborious and error-prone.  The
labeler would first have to understand what exactly each topic is capturing, in
itself a difficult task, then go through the entire data collection to answer
the question herself and determine how well the topic matches.  Clearly such a
process for creating a set of evaluation data is not feasible.

But what if the topic model fails to capture topics that truly answer the
question?  
Separate topics can be overlapping in words, where a human would consider them
the same topic.

The topic model itself could be high or low quality, and that affects the
results.

@@

TODO

>>>>

\begin{abstract}

Topic modeling is a popular and effective method for gathering meaning from a
set of documents.  Such models contain a lot of information and can be very
difficult to analyze.  For the proper visualization of these models, one must
present only the topics that are relevant to the task at hand.  We present the
novel task of topic ranking in response to user questions that greatly aids in
the visualization of topic models.  We make use of the document distribution of
a topic in order to rank the topics according to specific user questions.

\end{abstract}

@@

The rest of this paper is outlined as follows.  We first give further
motivation and a more detailed description of the task of ranking topics by
relevance in \secref{ranking}.  \secref{docdist} describes a method for
determining the relevance of a topic to various possible questions a person
could have when looking at a corpus, and \secref{examples} shows examples of
how this technique provides a proper ranking.  Finally, in \secref{conclusion}
we conclude.

@@

Our use of the document distribution for a topic focuses primarily on
calculating the entropy of the distribution and of related distributions where
documents are grouped by category.  For instance, one can group the documents
by author and compute the distribution of the topic over authors as
\begin{align}
  \label{eq:attrentropy}
  p(a|z) = \frac{\sum_{d \in \mathcal{A}} C(d,z)}{\sum_{d \in \mathcal{D}}
  C(d,z)}
\end{align}
where $a$ is an author and $\mathcal{A}$ is the set of documents written by
that author.

@@

\section{Example topic rankings}
\label{sec:examples}

We present here a few examples of questions a user might ask about a corpus and
how we can rank topics according to their relevance to those questions.  In our
examples we use a corpus of 480 campaign speeches from the 2008 presidential
primary and general election.  We inferred 150 topics in the data from a Gibbs
sampling of the LDA topic model.

The first question we address is, What topics are most prevelant in the corpus?
This is a simple question with an obvious answer that does not require using
$p(d|z)$.  One can simply order the topics by the number of tokens in the
topic.  Upon doing so one finds the following topics (we show the topic number
and the top ten words in the topic for the top five topics):

\begin{tabular}{l|l}
  35 & years time made long today place home day ago back \\
  42 & people don ve make country ll time put didn money \\
  119 & make national work effort including protect adiministration clear
	support policy \\
  15 & america world american future challenges great lead generation nation
	leadership \\
  108 & president work day senator house make white clinton care support \\
\end{tabular}

One can see that these topics for the most part are meaningless, capturing stop
words and noise in the data and hardly answering the question at hand.  When we
filter by a minimum value of coherence, however, we see a better picture:

\begin{tabular}{l|l}
  42 & people don ve make country ll time put didn money \\
  119 & make national work effort including protect adiministration clear
	support policy \\
  120 & law government rights respect case society fact laws principles order \\
  143 & jobs ve create economy make years billion companies invest energy \\
  101 & children family day man mother care love father parents young \\
\end{tabular}

While still somewhat noisy, this list gives the viewer some idea of what was
spoken of most during the 2008 presidential election.

Next we look at a similar question: What topics did John McCain talk about the
most?  This question is also answered without needing to use $p(d|z)$; simply
order the topics by the number of tokens that occur in speeches given by
McCain.  The list is the following (here again we filter by coherence):

\begin{tabular}{l|l}
  129 & mccain news john senator obama remarks america events donate email \\
  119 & make national work effort including protect adiministration clear
	support policy \\
  27 & mccain news john search media speeches www johnmccain email freedom \\
  90 & war iraq honor defeat enemy greater responsibility responsibilities
	difficult great \\
  66 & dollars reform federal washington congress reforms public policy
	problems system \\
\end{tabular}

Another similar, though slightly distinct, question is, What topics did Barack
Obama speak of most consistently throughout his campaign?  This question
differs from the previous in that the previous question could be answered by a
topic Obama used just a few times in long speeches.  This question looks for
topics that were consistent in all of his speeches.  Here we turn to the
document entropy for each topic as in \eqref{eq:entropy}, looking only at
speeches given by Obama.  The result of such an ordering is:

\begin{tabular}{l|l}
  60 &  barack remarks senator barackobama obama ll www http hope states\\
  47 & mccain senator john ll bush washington states election won george \\
  124 & change politics country time american washington ve common americans
	end \\
  84 & delegates south future party carolina kennedy support pledged total
	versus \\
  143 & jobs ve create economy make years billion companies invest energy \\
\end{tabular}

The first topic consists of markup on all of Barack Obama's speeches.  From the
rest of the topics one can see that Obama spoke a lot about his opponent,
Senator McCain, about change and the economy, and he apparently spent a lot of
time in South Carolina.

Another question one may have about these speeches is, What topics are most
consistent across candidates?  Each candidate has issues particular to them,
but what topics are universally used by the candidates?  A natural ranking for
this question is to order by entropy over candidates, as in
\eqref{eq:attrentropy}.  Such a ranking shows these to be universal topics:

\begin{tabular}{l|l}
  8 & congress government federal bill legislation members house conference
	statement congressional \\
  104 & states united countries international support economic state force
	military community \\
  103 & public political understand values home play rules shared committed
	important \\
  93 & president bush administration george war failed stop department
	september moral \\
  18 & result led group death allowed consequences early continued quickly
	balance \\
\end{tabular}

Similarly, one could ask, What topics were constant throughout the campaign?
This question is also answered with \eqref{eq:attrentropy}, with the month of
the speech used to group the documents together.  We see these topics as a
result:

\begin{tabular}{l|l}
  120 & law government rights respect case society fact laws principles order \\
  8 & congress government federal bill legislation members house conference
	statement congressional \\
  25 & party vote election campaign votes people democrats state democratic
	candidate \\
  101 & children family day man mother care love father parents young \\
  38 & economy economic american world security prosperity trade market markets
	growth \\
\end{tabular}

@@

Displaying the right topics to a user is essential to the effective
visualization of topic models.  We have presented the novel task of ranking
topics according to relevance to certain user questions.  We have also given a
method for ranking topics in order to answer a few specific questions.  Such
rankings allow visualizations of topic models that present information to users
in much more effective ways than was previously possible.

While our methods are useful for a few user questions, it remains as future
work to devise a method for ranking topics in answer to arbitrary questions.
The evaluation of such rankings is also a difficult task that must be left to
future work.
